\newape
\begin{solution}
  Two solution have been proposed
  \begin{itemize}
    \item
      Let's show that if $f$ is not one way, then $\Pi$ is not existentially unforgeable.
      Let $\A$ be the inverter of $f$, we will build $\A'$ that builds an existential forgery with non-negligible probability.

      \begin{itemize}
        \item $\A'$ receives $pk = (y_0, y_1)$
        \item $\A'$ ask the signature of 0 and gets $\sigma$, he does not really care about it
        \item $\A'$ gives $y_1$ to $\A$ which outputs $x_1'$
        \item $\A'$ outputs $(1, x_1')$
      \end{itemize}
      Since $y_1$ is the image of a random $x_1$, we are exactly in the inverting experiment so $f(x_1') = y_1$
      whith probability $\Pr[\Invert_{\A,f}(n) = 1]$.

      We know that
      \[
        \Pr[\Sigforge_{\A',\Pi}^{\text{1--time}}(n) = 1] = \Pr[\Invert_{\A,f}(n) = 1]
      \]
    \item
      Let assume that $f$ is not one way function $y = f(x)$.
      Then, we can recover $x$ with a non negligible probability $\epsilon_x (n)$.

      So, $(y_0, y_1) \Rightarrow (x_0, x_1)$ with probabilities $(\epsilon_{x_0} (n), \epsilon_{x_1} (n))$.
      We cannot compute a pre-image by asking the oracle. So I output $(0, x_1)$ and $(1, x_2)$ as a forgery.

      The probability $\Pr[\Sigforge(y)] = 2\epsilon(n)$, non-negligible since f is not one way.
  \end{itemize}
\end{solution}
\begin{solution}
  \begin{enumerate}
    \item
      A has $(i, \sigma(i))$ whith $\sigma (i) = f^{n-i} (x)$. We know (because $f$ is a permutation function) that:
      $$f(\sigma(i)) = f^{n-i+1}(x) = f^{n-(i-1)}(x) = \sigma(i-1)(x)$$
      Then it's possible to compute a valid forgery for every $j < i$. The scheme is then not a one time-signature.
    \item
      Need schema drawn at TP !! It's a lot simpler with it...

      $Pr[Success (A_\sigma)] = \epsilon(\lambda))$, $Pr[Abort] = \frac{n-k}{n}$, $Pr[Success] = \frac{n-k}{n-m-1}$ then:
      $$ Pr[Success(A_{owf})] = \epsilon(\lambda) \frac{n-k}{n} \frac{n-k}{n-m-1}$$

      If $\epsilon(\lambda)$ is not negligible, then the probability of success is not negligible.
    \item
      We have $s_k = (x, x')$, $p_k = (f^n(x), f^n(x'))$.
      Then $m \rightarrow \sigma = (f^{n-m}(x), f^m(x'))$.

  \end{enumerate}
\end{solution}
\begin{solution}
  \begin{enumerate}
    \item
      $(r, s) = ([g^k \pmod{p}] \pmod{q}, [H(m) + xr]k^{-1} \pmod{q})$, then:
      $u_1 = H(m)s^{-1} \pmod{q}$,  $u_2 = rs^{-1} \pmod{q}$, $r = [g^{u_1} g^{u_2} \pmod{p}] \pmod{q}$, $y = g^x$

      $$\Rightarrow g^{u_1 + xu_2} = g^{s^{-1}(H(m) + rx)} = [g^k \pmod{p}] \pmod{q} = r$$
    \item
      $s = (H(m) + xr)k^{-1} \pmod{q}$, $s' = (H(m') + xr)k^{-1} \pmod{q}$ ($s \neq s'$ otherwise we have a collision).
      $s - s' = (H(m) - H(m'))k^{-1} \pmod{q}$, $k = \frac{H(m) - H(m')}{s - s'}$.
      $s = (H(m) + xr)k^{-1}$ so $\frac{sk - H(m)}{r} = x$ where x is the secret.
  \end{enumerate}
\end{solution}

\nosolution

\copyapeexo{4}{1}
\copyapeexo{4}{2}
\copyapeexo{4}{3}
\copyapeexo{4}{4}
\copyapeexo{4}{5}
